---
title: "Time Series Project -- Sama & Mona"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(urca)
library(tseries)
```
1.Download the data set
```{r}
my_data<-read.csv("population dataset.csv")
my_data
```


```{r}
my_data<-my_data['value']
my_data<-my_data[-1,]
my_data<-round(my_data/1000000,2)
my_data
```

```{r}
my_data=ts(my_data, start = 1900)
plot.ts(my_data)
```

```{r}
df=ur.df(my_data,type='trend',lags=1)
summary(df)
```

```{r}
df.adf=adf.test(my_data,k=1)
df.adf
```



This dataset doesn't look stationary, try taking the difference.
```{r}
my_data_dif1<-diff(my_data,1)
plot.ts(my_data_dif1)
```

```{r}
print ('Lag = 1')
df_diff=ur.df(my_data_dif1,type='trend',lags=1)
summary(df_diff)

print("--------------------------------------")
df_adf1=adf.test(my_data_dif1,k=1)
df_adf1
```

```{r}
print ('Lag = 2')
df_diff=ur.df(my_data_dif1,type='trend',lags=2)
summary(df_diff)

print("--------------------------------------")
df_adf1=adf.test(my_data_dif1,k=2)
df_adf1
```

```{r}
print ('Lag = 3')
df_diff=ur.df(my_data_dif1,type='trend',lags=3)
summary(df_diff)

print("--------------------------------------")
df_adf1=adf.test(my_data_dif1,k=3)
df_adf1
```



```{r}
my_data_dif2<-diff(my_data,2)
plot.ts(my_data_dif2)
```



```{r}
df_diff2=ur.df(my_data_dif2,type='trend',lags=1)
summary(df_diff2)
```

```{r}
df_adf2=adf.test(my_data_dif2,k=1)
df_adf2
```


Plot the ACF and PACF.
```{r}
dif1_acf<-acf(my_data_dif1)
dif1_pacf<-pacf(my_data_dif1)
```
Therefore, the model seems to be arima(1,1,0)

Estimating the parameters (we'll try arima (1,1,0) and arima (0,1,1))

```{r}
ma1<-arima(my_data,order=c(0,1,1))
ma1
```



```{r}
ar1<-arima(my_data,order=c(1,1,0))
ar1
```
Model Becomes: 
X_t= phi*X_t-1+ epsilon
X_t=0.9885*X_t-1+0.0112

phi= 0.9885
epsilon= 0.0112

Checking the model assumptions, NICE assumptions
```{r}
par(mfrow=c(2,2))
qqnorm(ar1$residuals)
qqline(ar1$residuals)
plot.ts(ar1$residuals)
acf(ar1$residuals)
Box.test(ar1$residuals,lag=20,fitdf = 1)
```
Accept H0 -> errors are independent
Errors are normally distributed
Errors have an expectation equal to zero

```{r}
fr=forecast(ar1,h=5)
fr
```

Calculating the interval of each of our prediction intervals
```{r}
p=matrix(c(fr$lower[,2],fr$upper[,2],(fr$upper[,2]-fr$lower[,2])),ncol=3)
colnames(p)=c('Lower 95%','Upper 95%','Difference')
p
```


We forecasted the next 5 observation and observed the prediction interval at a 80% and 95% confidence level
point forecasts are in millions


